---
title: "Interpretation of the decidual immune microenvironment in RPL patients"
output:
  html_document:
    theme: united
  pdf_document: default
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---

```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  fig.width = 10,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  time_it = TRUE
)
```

In this tutorial, we utilized `ProjetSVR` to project decidual immune cells from both healthy individuals and patients with recurrent pregnancy loss (RPL) [<sup>1</sup>](#refer-anchor-1) onto a previously reported human maternal-fetal interface atlas [<sup>2</sup>](#refer-anchor-2).

## Download Reference Models and Query Data

```{r load_packages, results='hide'}
library(Seurat)
library(ProjectSVR)
library(tidyverse)

options(timeout = max(3600, getOption("timeout")))
```

```{r download_dataset, eval=FALSE}
# reference model
if (!dir.exists("models")) dir.create("models")
download.file(url = "https://zenodo.org/record/8350732/files/model.Vento2018.MFI.rds", 
              destfile = "models/model.Vento2018.MFI.rds")
# query data
if (!dir.exists("query")) dir.create("query")
download.file(url = "https://zenodo.org/record/8350748/files/Guo2021.MFI.seurat.slim.qs", 
              destfile = "query/Guo2021.MFI.seurat.slim.qs")
```

## Map Query to Reference

### Reference mapping

```{r reference_mapping, fig.width=4, fig.height=3}
reference <- readRDS("models/model.Vento2018.MFI.rds")
seu.q <- qs::qread("query/Guo2021.MFI.seurat.slim.qs")

seu.q$sample_id <- seu.q$orig.ident
seu.q$group <- factor(sub("[0-9]", "", seu.q$orig.ident), levels = c("Ctrl", "RPL"))

seu.q <- ProjectSVR::MapQuery(seu.q, reference = reference, add.map.qual = T, ncores = 10)

DimPlot(seu.q, reduction = "ref.umap", group.by = "sample_id", pt.size = .1)
```

### Maping quality

```{r map_quality, fig.width=14, fig.height=4}
## cutoff by adjusted p value
MapQCPlot(seu.q, map.q.cutoff = 2)
```

Visualize the projected query cells onto the reference atlas.

```{r projection_plot, fig.width=12, fig.height=5}
PlotProjection(seu.q, reference, split.by = "group", ref.color.by = "cell_type", 
               ref.size = .5, ref.alpha = .3, query.size = 1, query.alpha = .5, 
               n.row = 1, legend.ncol = 2)
```

### Label transfer

```{r label_transfer, fig.width=10, fig.height=4}
seu.q <- subset(seu.q, mean.knn.dist < 2)
seu.q <- ProjectSVR::LabelTransfer(seu.q, reference, ref.label.col = "cell_type")

DimPlot(seu.q, group.by = "knn.pred.celltype", split.by = "group") + 
  scale_color_manual(values = reference$ref.cellmeta$colors)
```

When we focus on the dNK population, we found that the dNK cells in RPL were shifted to dNK3 population, as reported in the original paper [<sup>1</sup>](#refer-anchor-1).

```{r projection_plot_dNK, fig.width=12, fig.height=5}
seu.q.dNK <- subset(seu.q, knn.pred.celltype %in% c("dNK p", "dNK1", "dNK2", "dNK3") )

PlotProjection(seu.q.dNK, reference, split.by = "group", ref.color.by = "cell_type", 
               ref.size = .5, ref.alpha = .3, query.size = .1, query.alpha = .1, 
               n.row = 1, legend.ncol = 2)
```

### Alluvia plot

```{r alluvia_plot, fig.width=4, fig.height=3}
AlluviaPlot(seu.q.dNK@meta.data, by = "group", 
            fill = "knn.pred.celltype",
            bar.width = .5, legend.ncol = 1)
```

### Compare the predicted labels vs the mannually annotated labels

```{r compare_heatmap, fig.width=10, fig.height=5}
data.stat <- table(seu.q$cell_type, seu.q$knn.pred.celltype)
data.stat <- data.stat[, colSums(data.stat) > 0]

pheatmap::pheatmap(data.stat, display_numbers = T, number_format = "%.0f", 
                   cluster_rows = F, cluster_cols = F, number_color = "black")
```


<details>
  <summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>


## References

<div id="refer-anchor-1"></div>
- [1] [Guo C, et al. Single-cell profiling of the human decidual immune microenvironment in patients with recurrent pregnancy loss. Cell Discov. 2021 Jan 4;7(1):1. doi: 10.1038/s41421-020-00236-z. PMID: 33390590.](https://pubmed.ncbi.nlm.nih.gov/33390590/)

<div id="refer-anchor-2"></div>
- [2] [Vento-Tormo R, et al. Single-cell reconstruction of the early maternal-fetal interface in humans. Nature. 2018 Nov;563(7731):347-353. doi: 10.1038/s41586-018-0698-6. Epub 2018 Nov 14. PMID: 30429548.](https://pubmed.ncbi.nlm.nih.gov/30429548/)

