---
title: "Project the perturbed germ cells onto mouse testicular cell atlas"
output:
  html_document:
    theme: united
  pdf_document: default
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---


```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  fig.width = 10,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  time_it = TRUE
)
```

To illustrate ProjectSVR's ability to enable the interpretation of defects of perturbed stages in a continuous developmental trajectory. Here, we provide a tutorial about how to project Zfp541-KO and WT germ cells [<sup>1</sup>](#refer-anchor-1) onto mouse testicular cell atlas using a pre-build model.

In this tutorial, we do not utilize the advanced wrapper functions (`MapQuery` and `LabelTransfer`) to illustrate how to use the low level functions.


## Download Related Dataset

```{r load_packages, results='hide'}
library(Seurat)
library(ProjectSVR)
library(tidyverse)

options(timeout = max(3600, getOption("timeout")))
```

```{r download_dataset, eval=FALSE}
# reference model
if (!dir.exists("models")) dir.create("models")
download.file(url = "https://zenodo.org/record/8350732/files/model.mTCA.rds", 
              destfile = "models/model.mTCA.rds")
# query data
if (!dir.exists("query")) dir.create("query")
download.file(url = "https://zenodo.org/record/8350748/files/query_Zfp541-KO.seurat.slim.qs", 
              destfile = "query/query_Zfp541-KO.seurat.slim.qs")
```

## Map Query to Reference

### Reference mapping

```{r reference_mapping, fig.width=12, fig.height=5}
reference <- readRDS("models/model.mTCA.rds")
seu.q <- qs::qread("query/query_Zfp541-KO.seurat.slim.qs")

genotype <- c("WT", "Zfp541-KO")
names(genotype) <- c("WT", "Z541")
seu.q$genotype <- factor(genotype[seu.q$orig.ident], levels = genotype)

## transform the counts matrix to gene set score matrix for query data
reference$gss.method # The reference model utilizes `AUCell` for gene set scoring.
top.genes <- reference$genes$gene.sets
gss.mat <- ComputeModuleScore(seu.q[["RNA"]]@data, gene.sets = top.genes, method = "AUCell", cores = 10)

## Project query into reference's UMAP spaces using gene set score matrix.
proj.res <- ProjectNewdata(feature.mat = gss.mat, model = reference$models$umap, cores = 10)
## Returning a `CellProject` object.
proj.res
## write the projected embeddings to original seurat object
seu.q[["ref.umap"]] <- CreateDimReducObject(proj.res@embeddings, key = "refUMAP_", assay = "RNA")

## visualization
p1 <- DimPlot(seu.q, reduction = "ref.umap", group.by = "genotype")
p2 <- DimPlot(seu.q, reduction = "ref.umap", group.by = "annotation", label = T) # original cell labels
(p1 + p2) & ggsci::scale_color_d3("category20")
```

### Mapping quality

```{r map_qc_threshold, fig.width=5, fig.height=3}
## calculate the map quanlity metrics (mean.knn.dist)
proj.res <- AddProjQual(object = proj.res, k = 20, repeats = 1e4)
head(proj.res@cellmeta)

## store the results in original seurat object
seu.q$mean.knn.dist <- proj.res@cellmeta$mean.knn.dist
seu.q$mapQ.p.val <- proj.res@cellmeta$p.val
seu.q$mapQ.p.adj <- proj.res@cellmeta$p.adj

## visualization
data.plot <- FetchData(seu.q, vars = c(paste0("refUMAP_", 1:2), "mean.knn.dist", "mapQ.p.adj"))

ggplot(data.plot, aes(mean.knn.dist, -log10(mapQ.p.adj))) +
  geom_point(size = .3) + 
  geom_vline(xintercept = 2.4, linetype = "dashed", color = "blue")
```

```{r map_qc_plot, fig.width=14, fig.height=4}
## cutoff by adjusted p value
MapQCPlot(seu.q, p.adj.cutoff = 1e-3)
## or mean.knn.dist
MapQCPlot(seu.q, map.q.cutoff = 2.4)
```

### Label transfer

```{r label_transfer, fig.width=8, fig.height=4}
## drop the low-quality mapped cells
seu.q <- subset(seu.q, mean.knn.dist < 2.4)

## input for KNN label transfer
ref.cellmeta <- reference$ref.cellmeta$meta.data 
# reference cell embeddings
ref.emb <- ref.cellmeta[, paste0("UMAP_", 1:2)]
# reference cell labels
ref.labels <- ref.cellmeta[["cell_type"]]
names(ref.labels) <- rownames(ref.cellmeta)
# query cell embeddings
query.emb <- seu.q[["ref.umap"]]@cell.embeddings

## KNN label transfer
knn.pred.res <- KnnLabelTransfer(query.emb = query.emb, ref.emb = ref.emb, 
                                 ref.labels = ref.labels, k = 100)
## over-cluster and then performing majority voting for each clusters
knn.pred.mv <- MajorityVote(feature.mat = gss.mat, 
                            cell.types = knn.pred.res[, c("labels"), drop = F], 
                            k = 100, min.prop = 0.3)

## save results to seurat object
seu.q$knn.pred.celltype <- knn.pred.res$labels
seu.q$knn.pred.celltype.major_votes <- knn.pred.mv$labels.major_votes
ref.celltype.levels <- levels(ref.cellmeta[["cell_type"]])
seu.q$knn.pred.celltype <- factor(seu.q$knn.pred.celltype, levels = ref.celltype.levels)
seu.q$knn.pred.celltype.major_votes <- factor(seu.q$knn.pred.celltype.major_votes, 
                                              levels = ref.celltype.levels)

DimPlot(seu.q, reduction = "ref.umap", group.by = "knn.pred.celltype.major_votes", 
        split.by = "genotype") + ggsci::scale_color_d3()
```

```{r compare_heatmap, fig.width=6, fig.height=5}
data.stat <- table(seu.q$annotation, seu.q$knn.pred.celltype.major_votes)
data.stat <- data.stat[, colSums(data.stat) > 0]

pheatmap::pheatmap(data.stat, display_numbers = T, number_format = "%.0f", 
                   cluster_rows = F, cluster_cols = F,
                   number_color = "black")
```

<details>
  <summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>

## References

<div id="refer-anchor-1"></div>
- [1] [Xu J, et al. ZFP541 maintains the repression of pre-pachytene transcriptional programs and promotes male meiosis progression. Cell Rep. 2022 Mar 22;38(12):110540. doi: 10.1016/j.celrep.2022.110540. PMID: 35320728.](https://pubmed.ncbi.nlm.nih.gov/35320728/)
