---
title: "Project the NRRA induced meiotic cells onto mTCA"
output:
  html_document:
    theme: united
  pdf_document: default
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---

```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  fig.width = 10,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  time_it = TRUE
)
```

Adopting the organ-scale reference atlas to evaluate the in vitro development procedure is an appealing application. In this tutorial, we projected germ cells that were dissociated from the testis of 6-day-old mouse pups and undergoing in vitro meiosis induced by co-treatment of nutrient restriction and retinoic acid (NRRA) [<sup>1</sup>](#refer-anchor-1) onto mouse testicular cell atlas (mTCA) using a pre-build model.

## Download Related Dataset

```{r load_packages, results='hide'}
library(Seurat)
library(ProjectSVR)
library(tidyverse)

options(timeout = max(3600, getOption("timeout")))
```


```{r download_dataset, eval=FALSE}
# reference model
if (!dir.exists("models")) dir.create("models")
download.file(url = "https://zenodo.org/record/8350732/files/model.mTCA.rds", 
              destfile = "models/model.mTCA.rds")
# query data
if (!dir.exists("query")) dir.create("query")
download.file(url = "https://zenodo.org/record/8350748/files/query_NRRA_induced_meiotic_cells.seurat.slim.qs", 
              destfile = "query/query_NRRA_induced_meiotic_cells.seurat.slim.qs")
```

## Map Query to Reference

```{r reference_mapping, fig.width=4, fig.height=3}
reference <- readRDS("models/model.mTCA.rds")
seu.q <- qs::qread("query/query_NRRA_induced_meiotic_cells.seurat.slim.qs")

seu.q <- ProjectSVR::MapQuery(seu.q, reference = reference, add.map.qual = T, ncores = 10)

DimPlot(seu.q, reduction = "ref.umap", group.by = "day")
```

### Maping quality

```{r map_quality, fig.width=14, fig.height=4}
## cutoff by adjusted p value
MapQCPlot(seu.q, p.adj.cutoff = 0.1)
```

Visualize the projected query cells onto the reference atlas.

```{r projection_plot, fig.width=15, fig.height=12}
PlotProjection(seu.q, reference, split.by = "day", ref.color.by = "cell_type", 
               ref.size = .5, ref.alpha = .3, query.size = 1, query.alpha = .5, 
               n.row = 2, legend.ncol = 1)
```


### Label transfer

```{r label_transfer, fig.width=14, fig.height=4}
seu.q <- subset(seu.q, mapQ.p.adj < 0.1)
seu.q <- ProjectSVR::LabelTransfer(seu.q, reference, ref.label.col = "cell_type")
## Majority votes
feature.mat <- FetchData(seu.q, vars = rownames(seu.q))
cell.types <- FetchData(seu.q, vars = c("knn.pred.celltype"))
knn.pred.mv <- MajorityVote(feature.mat = feature.mat, cell.types = cell.types, k = 100, min.prop = 0.3)
seu.q$knn.pred.celltype.major_votes <- knn.pred.mv$knn.pred.celltype.major_votes

DimPlot(seu.q, group.by = "knn.pred.celltype.major_votes", split.by = "day") + 
  ggsci::scale_color_d3()
```

Note: The `m9-PGC_Piwil4_arrest` is a wrong label because its location is on the center of m9, m26, and m33.

We validated this by checking the pan somatic marker `Vim` and pan germ cell marker `Ddx4`. And we found `m9` is not germ cell, but also expressed lower `Vim`.

This observation remind us that we should double check the label transfer results.

```{r, fig.width=8, fig.height=4}
DefaultAssay(seu.q) <- "RNA"
seu.q[["RNA"]]@counts <- seu.q[["RNA"]]@data
seu.q <- NormalizeData(seu.q)
FeaturePlot(seu.q, reduction = "ref.umap", features = c("Vim", "Ddx4"), pt.size = .5)
```

I am quite curious about the identity of m9. So I calculated the m9 specific expressed genes.

```{r, fig.width=8, fig.height=4}
Idents(seu.q) <- seu.q$knn.pred.celltype.major_votes
seu.q.ds <- subset(seu.q, downsample = 200)
m9.markers <- FindMarkers(seu.q.ds, ident.1 = "m9-PGC_Piwil4_arrest", only.pos = T)
head(m9.markers, n = 10)
FeaturePlot(seu.q, reduction = "ref.umap", features = c("Foxi3", "Prr23a1"), pt.size = .5)
```

We found:

1. The "m9" markers are not correlated to any cell type in the testicular cell atlas. 

2. The "m9" do not exist in day0 group.

These suggest that "m9" may be a novel cell type induced by NRRA in vitro.

```{r}
# devtools::install_github("JarningGau/scutilsR")
library(scutilsR)
data("mca_mmu")
data("hcl_mmu")
t2g <- rbind(mca_mmu, hcl_mmu)

genes <- subset(m9.markers, p_val_adj < 1e-6 & avg_log2FC > 1) %>% rownames()
e.res <- clusterProfiler::enricher(genes, TERM2GENE = t2g)
head(e.res@result)
```

According to the enriched terms of cell markers, the m9 may be a type of endothelial cells.

### Alluvia plot

Then we focus on the NRRA induced meiosis progression.

```{r alluvia_plot, fig.width=6, fig.height=5}
cellmeta <- subset(seu.q@meta.data, grepl("m1[12345]+", knn.pred.celltype.major_votes))

AlluviaPlot(cellmeta, by = "day", 
            fill = "knn.pred.celltype.major_votes",
            colors = reference$ref.cellmeta$colors,
            bar.width = .5, legend.ncol = 1)
```

We found 4~5% germ cells entering meiosis after 3 days treatment by NRRA.


<details>
<summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>


## References

<div id="refer-anchor-1"></div>
- [1] [Zhang X, et al. Nutrient restriction synergizes with retinoic acid to induce mammalian meiotic initiation in vitro. Nat Commun. 2021 Mar 19;12(1):1758. doi: 10.1038/s41467-021-22021-6. PMID: 33741948](https://pubmed.ncbi.nlm.nih.gov/33741948/)

