---
title: "Training the ensemble SVM model to predict cell type directly"
output:
  html_document:
    theme: united
  pdf_document: default
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---


```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  #tidy = TRUE,
  #tidy.opts = list(width.cutoff = 95),
  fig.width = 10,
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

ProjectSVR also implemented a wrapper `FitEnsemblMultiClassif()` to train an ensemble SVM model for cell type auto-annotation. In this tutorial, we show how to train such a model and use it for cell type annotation.

## Download Related Dataset

```{r load_packages, results='hide'}
library(ProjectSVR)
library(Seurat)
library(tidyverse)
options(timeout = max(3600, getOption("timeout")))
`%notin%` <- Negate(`%in%`)
```

```{r download_dataset, eval=FALSE}
if (!dir.exists("models")) dir.create("models")
if (!dir.exists("reference")) dir.create("reference")
if (!dir.exists("query")) dir.create("query")

# reference model
download.file(url = "https://zenodo.org/record/8350732/files/model.disco_pbmc.rds", 
              destfile = "models/model.disco_pbmc.rds")

# reference data
download.file(url = "https://zenodo.org/record/8350746/files/mTCA.seurat.slim.qs", 
              destfile = "reference/DISCO_hPBMCs.seurat.slim.qs")
# query data
download.file(url = "https://zenodo.org/record/8350748/files/query_hPBMCs.seurat.slim.qs", 
              destfile = "query/query_hPBMCs.seurat.slim.qs")
```

## Build Reference Model

```{r load_reference, fig.width=9, fig.height=5}
data("pals")
seu.ref <- qs::qread("reference/DISCO_hPBMCs.seurat.slim.qs")
p1 <- DimPlot(seu.ref, pt.size = .4) + scale_color_manual(values = pals$disco_blood)
LabelClusters(p1, id = "ident")
```

### Transfer raw count matrix to gene set score matrix

```{r add_gene_module_score}
reference <- readRDS("models/model.disco_pbmc.rds")
top.genes <- reference$genes$gene.sets
bg.genes <- reference$genes$bg.genes
reference$gss.method

seu.ref <- ComputeModuleScore(seu.ref, gene.sets = top.genes, bg.genes = bg.genes, method = "UCell", cores = 5)
# The signature score matrix is stored in 'SignatureScore' assay
Assays(seu.ref)
DefaultAssay(seu.ref) <- "SignatureScore"
```

### Training reference model

```{r train_model}
gss.mat <- FetchData(seu.ref, vars = rownames(seu.ref))
cell.types <- FetchData(seu.ref, vars = c("cell_type", "cell_subtype"))
batch.size = 8000 # number of subsampled cells for each SVR model 
n.models = 20      # number of SVR models trained
svm.model <- FitEnsemblMultiClassif(feature.mat = gss.mat, 
                                    cell.types = cell.types,
                                    batch.size = batch.size,
                                    n.models = n.models, 
                                    balance.cell.type = TRUE, # balanced sampling for each cell label
                                    cores = 10)
```

```{r save_model}
## save model to reference object
reference$models$cell_type <- svm.model

qs::qsave(reference, "models/model.disco_pbmc.v2.qs")
```

## Cell Type Annotation

```{r auto_annotation, fig.width=12, fig.height=5}
seu.q <- qs::qread("query/query_hPBMCs.seurat.slim.qs")

## map query
seu.q <- ProjectSVR::MapQuery(seu.q, reference = reference, add.map.qual = T, ncores = 10)
seu.q

## predict cell type
gss.mat.q <- FetchData(seu.q, vars = rownames(seu.q))
pred.res <- PredictNewdata(feature.mat = gss.mat.q, model = svm.model, cores = 10)
head(pred.res)

## save results to seurat object
seu.q$cell_type.pred <- pred.res$cell_type
seu.q$cell_subtype.pred <- pred.res$cell_subtype

## visualization
p1 <- DimPlot(seu.q, reduction = "ref.umap", group.by = c("cell_type")) + ggsci::scale_color_d3("category20")
p2 <- DimPlot(seu.q, reduction = "ref.umap", group.by = c("cell_type.pred")) + ggsci::scale_color_d3()

p1 <- LabelClusters(p1, id = "cell_type")
p2 <- LabelClusters(p2, id = "cell_type.pred")

p1 + p2
```

```{r, visualization, fig.width=14, fig.height=5}
p1 <- DimPlot(seu.q, reduction = "ref.umap", group.by = c("cell_subtype")) + ggsci::scale_color_d3("category20")
p2 <- DimPlot(seu.q, reduction = "ref.umap", group.by = c("cell_subtype.pred")) + scale_color_manual(values = pals$disco_blood)

p1 <- LabelClusters(p1, id = "cell_subtype")
p2 <- LabelClusters(p2, id = "cell_subtype.pred")

p1 + p2
```

The ProjectSVR gives pretty good cell type predictions. It is quite interesting that the ProjectSVR predicts GZMK+ NK cells. To verify this, we check a marker combination and found that GZMK+ NK cells can be defined by NKG7+/GZMK+/CD16-/CD8A-, which supports the ProjectSVR's prediction.

```{r markers, fig.width=10, fig.height=8}
DefaultAssay(seu.q) <- "RNA"
seu.q[["RNA"]]@counts <- seu.q[["RNA"]]@data
seu.q <- NormalizeData(seu.q)
FeaturePlot(seu.q, reduction = "ref.umap", features = c("GZMK", "FCGR3A", "NKG7", "CD8A"), ncol = 2)
```


<details>
<summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>


