---
title: "Build reference model for pan cancer T cell lanscape"
output:
  html_document:
    theme: united
  pdf_document: default
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---


```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  # tidy = TRUE,
  # tidy.opts = list(width.cutoff = 95),
  fig.width = 10,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  time_it = TRUE
)
```

In this tutorial, we showed how to build reference model for a tumor infiltrated pan-cancer T cell atlas [<sup>1</sup>](#refer-anchor-1).

## Download Related Dataset

```{r load_packages, results='hide'}
library(Seurat)
library(ProjectSVR)
library(tidyverse)
options(timeout = max(3600, getOption("timeout")))
```

```{r download_dataset, eval=FALSE}
if (!dir.exists("reference")) dir.create("reference")

download.file(url = "https://zenodo.org/record/8350746/files/ZhengLiangtao.CD4.seurat.slim.qs", 
              destfile = "reference/ZhengLiangtao.CD4.seurat.slim.qs")

download.file(url = "https://zenodo.org/record/8350746/files/ZhengLiangtao.CD8.seurat.slim.qs", 
              destfile = "reference/ZhengLiangtao.CD8.seurat.slim.qs")
```

## Reference Model for Pan-cancer CD4 T Cell Landscape

The original paper [<sup>1</sup>](#refer-anchor-1) adopted a "metacell-like" strategy for data integration. We mapped the coordinates of metacells in UMAP space to single cells and trained supported vector regression (SVR) model on them.

```{r referenc_plot, fig.width=8, fig.height=4}
data("pals")
seu.ref <- qs::qread("reference/ZhengLiangtao.CD4.seurat.slim.qs")

data.plot <- seu.ref@misc$data.refplot$meta.data

text.pos <- data.frame(
  x = c(-4.5, 4, 3.5,  2, -4.5, -7, -6),
  y = c(  5,  4, -4,  -6, -6,   -2,  2),
  label = c("Tn", "Treg", "Tfh", "Tfh/Th1", "Temra", "Tem", "Tm")
)

## meta cell reference
ggplot(data.plot, aes(UMAP_1, UMAP_2, color = cluster.name), pt.size = .4) + 
  geom_point(size = .3) + 
  ggtitle("Pan-cancer CD4 T cell") + 
  scale_color_manual(values = pals$`pan-cancer_cd4t`) +
  geom_text(inherit.aes = F, data = text.pos, mapping = aes(x, y, label = label), size = 4) + 
  guides(color = guide_legend(override.aes = list(size = 2))) + 
  coord_equal(ratio = 1) + 
  theme(legend.text = element_text(size = 7),
        plot.title = element_text(hjust = .5))
```

### Transfer raw count matrix to gene set score matrix

We utilized the author defined markers stored in `seu.ref@misc$markers` as features to calculate the gene set scores. These markers were ordered by significance. We extracted the top 20 most significant genes for each cell state.

```{r add_gene_module_score, fig.width=8, fig.height=4}
top.genes <- seu.ref@misc$markers
top.genes <- lapply(top.genes, function(xx) head(xx, 20))
sapply(top.genes, length)

seu.ref <- ComputeModuleScore(seu.ref, gene.sets = top.genes, method = "UCell", cores = 5)

# The signature score matrix is stored in 'SignatureScore' assay
Assays(seu.ref)
DefaultAssay(seu.ref) <- "SignatureScore"
FeaturePlot(seu.ref, features = c("CD4.c12", "CD4.c16"), pt.size = 3)
```

### Training reference model

```{r train_model}
gss.mat <- FetchData(seu.ref, vars = rownames(seu.ref))
embeddings.df <- FetchData(seu.ref, vars = paste0("UMAP_", 1:2))
batch.size = 8000 # number of subsampled cells for each SVR model 
n.models = 20      # number of SVR models trained
umap.model <- FitEnsembleSVM(feature.mat = gss.mat,
                             emb.mat = embeddings.df,
                             batch.size = batch.size,
                             n.models = n.models,
                             cores = 10)
```

### Create reference object

1. meta.data: metacell meta data (embeddings & cell type information)

2. gss.method: method used in `ComputeModuleScore()`

3. [optional] colors: for plots

4. [optional] text.pos: text annotation on the reference plots

```{r create_reference}
meta.data <- seu.ref@misc$data.refplot$meta.data
colors <- pals$`pan-cancer_cd4t`
bg.genes <- rownames(seu.ref[["RNA"]])

reference <- CreateReference(umap.model = umap.model, 
                             gene.sets = top.genes, 
                             bg.genes = bg.genes,
                             meta.data = meta.data, 
                             gss.method = "UCell",
                             colors = colors, 
                             text.pos = text.pos)

dir.create("models")
saveRDS(reference, "models/model.Zheng2021.CD4Tcell.rds")
```

## Reference Model for Pan-cancer CD8 T Cell Landscape

```{r referenc_plot2, fig.width=6, fig.height=5}
data("pals")
seu.ref <- qs::qread("reference/ZhengLiangtao.CD8.seurat.slim.qs")
gc()

data.plot <- seu.ref@misc$data.refplot$meta.data

text.pos <- data.frame(
  x = c(-6, -5, 2, 6,  3, -5, -4.5, 0),
  y = c( 1,  7, 6, 5, -7, -5, -1  , 0),
  label = c("ISG+", "Tex", "KIR+ NK-like", "Temra", "Tn", "Tc17", "Trm", "Tem")
)

## meta cell reference
ggplot(data.plot, aes(UMAP_1, UMAP_2, color = cluster.name), pt.size = .4) + 
  geom_point(size = .3) + 
  ggtitle("Pan-cancer CD8 T cell") + 
  scale_color_manual(values = pals$`pan-cancer_cd8t`) +
  geom_text(inherit.aes = F, data = text.pos, mapping = aes(x, y, label = label), size = 4) + 
  guides(color = guide_legend(override.aes = list(size = 2), ncol = 1)) + 
  coord_equal(ratio = 1) + 
  theme(legend.text = element_text(size = 7),
        plot.title = element_text(hjust = .5))
```

### Transfer raw count matrix to gene set score matrix

```{r add_gene_module_score2, fig.width=8, fig.height=4}
top.genes <- seu.ref@misc$markers
top.genes <- lapply(top.genes, function(xx) head(xx, 20))
sapply(top.genes, length)

seu.ref <- ComputeModuleScore(seu.ref, gene.sets = top.genes, method = "UCell", cores = 5)

# The signature score matrix is stored in 'SignatureScore' assay
Assays(seu.ref)
DefaultAssay(seu.ref) <- "SignatureScore"
FeaturePlot(seu.ref, features = c("CD8.c12", "CD8.c16"), pt.size = 3)
```

### Training reference model

```{r train_model2}
gss.mat <- FetchData(seu.ref, vars = rownames(seu.ref))
embeddings.df <- FetchData(seu.ref, vars = paste0("UMAP_", 1:2))
batch.size = 8000 # number of subsampled cells for each SVR model 
n.models = 20      # number of SVR models trained
umap.model <- FitEnsembleSVM(feature.mat = gss.mat,
                             emb.mat = embeddings.df,
                             batch.size = batch.size,
                             n.models = n.models,
                             cores = 10)
```

### Create reference object

```{r create_reference2}
meta.data <- seu.ref@misc$data.refplot$meta.data
colors <- pals$`pan-cancer_cd8t`
bg.genes <- rownames(seu.ref[["RNA"]])

reference <- CreateReference(umap.model = umap.model, 
                             gene.sets = top.genes, 
                             bg.genes = bg.genes,
                             meta.data = meta.data, 
                             gss.method = "UCell",
                             colors = colors, 
                             text.pos = text.pos)

dir.create("models")
saveRDS(reference, "models/model.Zheng2021.CD8Tcell.rds")
```


<details>
  <summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>

## References

<div id="refer-anchor-1"></div>
- [1] [Zheng L, et al. Pan-cancer single-cell landscape of tumor-infiltrating T cells. Science. 2021 Dec 17;374(6574):abe6474. doi: 10.1126/science.abe6474. Epub 2021 Dec 17. PMID: 34914499.](https://pubmed.ncbi.nlm.nih.gov/34914499/)


