---
title: "Build reference model for DISCO blood atlas"
output:
  html_document:
    theme: united
  pdf_document: default
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---

```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  #tidy = TRUE,
  #tidy.opts = list(width.cutoff = 95),
  fig.width = 10,
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

In this tutorial, we showed how to build the reference model for blood atlas from [DISCO](https://www.immunesinglecell.org/) database. This case showed that once the reference atlas is well annotated and integrated into a harmonious low dimensional embeddings, we can easily training a supported vector regression model from gene set scores to representated embeddings via `ProjectSVR`.

## Download Reference Atlas

```{r load_packages, results='hide'}
library(ProjectSVR)
library(Seurat)
library(tidyverse)
options(timeout = max(3600, getOption("timeout")))
`%notin%` <- Negate(`%in%`)
```

```{r download_dataset, eval=FALSE}
if (!dir.exists("reference")) dir.create("reference")

download.file(url = "https://zenodo.org/record/8350746/files/DISCO_hPBMCs.seurat.slim.qs", 
              destfile = "reference/DISCO_hPBMCs.seurat.slim.qs")
```

## Build Reference Model

```{r load_reference, fig.width=9, fig.height=5}
data("pals")
seu.ref <- qs::qread("reference/DISCO_hPBMCs.seurat.slim.qs")
p1 <- DimPlot(seu.ref, pt.size = .4) + scale_color_manual(values = pals$disco_blood)
LabelClusters(p1, id = "ident")
```

### Extract signatures

Here we extract the top25 marker genes for each cell type (ribosomal and mitochondrial genes were removed).

```{r find_markers}
data(ribo.genes)

table(Idents(seu.ref))

## To accelerate the calculation, we downsampled 200 cells from each cluster
seu.ref.ds <- subset(seu.ref, downsample = 200)

## The raw count matrix stores in `data` slot, so we copy it to `counts` slot for normalization.
seu.ref.ds[["RNA"]]@counts <- seu.ref.ds[["RNA"]]@data
seu.ref.ds <- NormalizeData(seu.ref.ds)

## Parallel calculation of the cell markers.
all.markers <- mcFindAllMarkers(seu.ref.ds, do.flatten = F, n.cores = 10)

top.genes <- lapply(all.markers, function(xx){
  yy <- subset(xx, p_val_adj < 1e-6 & avg_log2FC > log2(1.5))
  yy <- subset(yy, Gene.name.uniq %notin% ribo.genes)
  yy <- yy[!grepl("^MT-", yy$Gene.name.uniq), ]
  head(yy$Gene.name.uniq, 25)
})

sapply(top.genes, length)
# rename the gene set
# [Note] Avoid to use '_' or '-' in the gene set names.
names(top.genes) <- paste0("feature.", 1:length(top.genes))
# the background genes are the union of all markers
bg.genes <- do.call(c, top.genes) %>% unique()
```

### Transfer raw count matrix to gene set score matrix

```{r add_gene_module_score}
seu.ref <- ComputeModuleScore(seu.ref, gene.sets = top.genes, bg.genes = bg.genes, 
                              method = "UCell", cores = 5)
# The signature score matrix is stored in 'SignatureScore' assay
Assays(seu.ref)
DefaultAssay(seu.ref) <- "SignatureScore"
```

### Training reference model

```{r train_model}
gss.mat <- FetchData(seu.ref, vars = rownames(seu.ref))
embeddings.df <- FetchData(seu.ref, vars = paste0("UMAP_", 1:2))
batch.size = 8000 # number of subsampled cells for each SVR model 
n.models = 20      # number of SVR models trained
umap.model <- FitEnsembleSVM(feature.mat = gss.mat,
                             emb.mat = embeddings.df,
                             batch.size = batch.size,
                             n.models = n.models,
                             cores = 10)
```

### Create reference object

1. meta.data: cell meta data (embeddings & cell type information)

2. gss.method: method used in `ComputeModuleScore()`

3. [optional] colors: for plots

4. [optional] text.pos: text annotation on the reference plots

```{r create_reference}
meta.data <- FetchData(seu.ref, vars = c(paste0("UMAP_", 1:2), "cell_type", "cell_subtype") )
colors <- pals$disco_blood
text.pos <- data.frame(
  x = c(11,  2, 6, -2, 0, 13, 10, 14,  7, -2,  1, 3, 4,   7),
  y = c(-4, -4, 1,  1, 5,  2,  4, 12, 16, 12, 15, 9, 6.5, 5),
  label = c("B", "pDC", "plasma", "cDC", "mono", "CD4 T", "Treg", "CD8 T", 
            "NK", "neutrophil", "platelet", "RBC", "HSC", "T/NK pro")
)

reference <- CreateReference(umap.model = umap.model, 
                             gene.sets = top.genes, 
                             bg.genes = bg.genes,
                             meta.data = meta.data, 
                             gss.method = "UCell",
                             colors = colors, 
                             text.pos = text.pos)

dir.create("models")
saveRDS(reference, "models/model.disco_pbmc.rds")
```


<details>
  <summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>
