---
title: "Build reference model for human maternal-fetal interface atlas"
output:
  html_document:
    theme: united
  pdf_document: default
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---

```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  #tidy = TRUE,
  #tidy.opts = list(width.cutoff = 95),
  fig.width = 10,
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

In this tutorial, we showed how to build reference model for human maternal-fetal interface cell atlas [<sup>1</sup>](#refer-anchor-1).

## Download Reference Atlas

```{r load_packages, results='hide'}
library(ProjectSVR)
library(Seurat)
library(tidyverse)
options(timeout = max(3600, getOption("timeout")))
`%notin%` <- Negate(`%in%`)
```

```{r download_dataset, eval=FALSE}
if (!dir.exists("reference")) dir.create("reference")

download.file(url = "https://zenodo.org/record/8350746/files/Vento2018.MFI.seurat.slim.qs", 
              destfile = "reference/Vento2018.MFI.seurat.slim.qs")
```

## Build Reference Model

```{r load_reference, fig.width=8, fig.height=5}
data("pals")
seu.ref <- qs::qread("reference/Vento2018.MFI.seurat.slim.qs")
p1 <- DimPlot(seu.ref, pt.size = .1) + scale_color_manual(values = pals$hfmi)
LabelClusters(p1, id = "ident")
```

### Extract signatures

Here we extract the top25 marker genes for each cell type (ribosomal and mitochondrial genes were removed).

```{r find_markers}
data(ribo.genes)

table(Idents(seu.ref))

## To accelerate the calculation, we downsampled 200 cells from each cluster
seu.ref.ds <- subset(seu.ref, downsample = 200)

## The raw count matrix stores in `data` slot, so we copy it to `counts` slot for normalization.
seu.ref.ds[["RNA"]]@counts <- seu.ref.ds[["RNA"]]@data
seu.ref.ds <- NormalizeData(seu.ref.ds)

## Parallel calculation of the cell markers.
all.markers <- mcFindAllMarkers(seu.ref.ds, do.flatten = F, n.cores = 10)

top.genes <- lapply(all.markers, function(xx){
  yy <- subset(xx, p_val_adj < 1e-6 & avg_log2FC > log2(1.5))
  yy <- subset(yy, Gene.name.uniq %notin% ribo.genes)
  yy <- yy[!grepl("^MT-", yy$Gene.name.uniq), ]
  head(yy$Gene.name.uniq, 25)
})

sapply(top.genes, length)
# rename the gene set
# [Note] Avoid to use '_' or '-' in the gene set names.
names(top.genes) <- paste0("feature.", 1:length(top.genes))
# the background genes are the union of all markers
bg.genes <- do.call(c, top.genes) %>% unique()
```

### Transfer raw count matrix to gene set score matrix

```{r add_gene_module_score}
seu.ref <- ComputeModuleScore(seu.ref, gene.sets = top.genes, bg.genes = bg.genes, 
                              method = "UCell", cores = 5)
# The signature score matrix is stored in 'SignatureScore' assay
Assays(seu.ref)
DefaultAssay(seu.ref) <- "SignatureScore"
```

### Training reference model

```{r train_model}
gss.mat <- FetchData(seu.ref, vars = rownames(seu.ref))
embeddings.df <- FetchData(seu.ref, vars = paste0("UMAP_", 1:2))
batch.size = 8000 # number of subsampled cells for each SVR model 
n.models = 20      # number of SVR models trained
umap.model <- FitEnsembleSVM(feature.mat = gss.mat,
                             emb.mat = embeddings.df,
                             batch.size = batch.size,
                             n.models = n.models,
                             cores = 10)
```

### Create reference object

1. meta.data: cell meta data (embeddings & cell type information)

2. gss.method: method used in `ComputeModuleScore()`

3. [optional] colors: for plots

4. [optional] text.pos: text annotation on the reference plots

```{r create_reference}
meta.data <- FetchData(seu.ref, vars = c(paste0("UMAP_", 1:2), "annotation"))
colnames(meta.data)[3] <- "cell_type"

colors <- pals$hfmi
text.pos <- data.frame(
  x = c(-13, -5, 7, -10, -2, 2, 5, 5, 10, 10, 5, -1,-3, 0, 5, 0, 10),
  y = c(-7, -11, -11, 2, -2,-4, 8, 12, 5, 10, 5,  9, 5, 3,-1, 1, -5),
  label = c("dNK", "Endo", "Stromal", "NK", "T cell", "Epi", "SCT", "VCT", "EVT", "Trophoblast", "HB", 
            "Macrophage", "DC", "Monocyte", "Fibroblast", "Granulocyte", "Perivascular")
)

reference <- CreateReference(umap.model = umap.model, 
                             gene.sets = top.genes, 
                             bg.genes = bg.genes,
                             meta.data = meta.data, 
                             gss.method = "UCell",
                             colors = colors, 
                             text.pos = text.pos)

dir.create("models")
saveRDS(reference, "models/model.Vento2018.MFI.rds")
```


<details>
  <summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>


## References

<div id="refer-anchor-1"></div>
- [1] [Vento-Tormo R, et al. Single-cell reconstruction of the early maternal-fetal interface in humans. Nature. 2018 Nov;563(7731):347-353. doi: 10.1038/s41586-018-0698-6. Epub 2018 Nov 14. PMID: 30429548.](https://pubmed.ncbi.nlm.nih.gov/30429548/)

