---
title: "Quick start of ProjectSVR"
output:
  html_document:
    theme: united
  pdf_document: default
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---

```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  fig.width = 10,
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

## Download related dataset

```{r, results='hide'}
library(ProjectSVR)
library(Seurat)
library(tidyverse)
options(timeout = max(3600, getOption("timeout")))
`%notin%` <- Negate(`%in%`)
```

```{r, eval=FALSE}
# reference datass
download.file(url = "https://zenodo.org/record/8147304/files/disco_pbmc_small.seurat.slim.rds", destfile = "disco_pbmc_small.seurat.slim.rds")
# query data
download.file(url = "https://zenodo.org/record/8147304/files/query_pbmc_small.seurat.slim.rds", destfile = "query_pbmc_small.seurat.slim.rds")
```


## Build Reference Model

```{r, fig.width=9, fig.height=5}
seu.ref <- readRDS("disco_pbmc_small.seurat.slim.rds")

DimPlot(seu.ref, pt.size = .4) +
  scale_color_manual(values = seu.ref@misc$data.refplot$colors) +
  geom_text(inherit.aes = F, data = seu.ref@misc$data.refplot$text.pos,
            mapping = aes(x, y, label = label), size = 4)
```

### Extract signatures

Here we extract the top25 marker genes for each cell type (ribosomal and mitochondrial genes were removed).

```{r}
data(ribo.genes)

table(Idents(seu.ref))
seu.ref[["RNA"]]@counts <- seu.ref[["RNA"]]@data
seu.ref <- NormalizeData(seu.ref)
all.markers <- mcFindAllMarkers(seu.ref, do.flatten = F, n.cores = 20)

top.genes <- lapply(all.markers, function(xx){
  yy <- subset(xx, p_val_adj < 1e-6 & avg_log2FC > log2(1.5))
  yy <- subset(yy, Gene.name.uniq %notin% ribo.genes)
  yy <- yy[!grepl("^MT-", yy$Gene.name.uniq), ]
  head(yy$Gene.name.uniq, 25)
})

sapply(top.genes, length)
names(top.genes) <- paste0("feature_", 1:length(top.genes))

bg.genes <- do.call(c, top.genes) %>% unique()
```

### Transfer raw count matrix to gene set score matrix

```{r}
seu.ref <- seu.ref[bg.genes, ]
seu.ref <- UCell::AddModuleScore_UCell(seu.ref, features = top.genes, ncores = 20)
gss.mat <- select(seu.ref@meta.data, ends_with("_UCell"))
```

### Training reference model

```{r}
embeddings.df <- FetchData(seu.ref, vars = paste0("UMAP_", 1:2))
batch.size = 4000 # number of subsampled cells for each SVR model 
n.models = 5      # number of SVR models trained
umap.model <- FitEnsembleSVM(feature.mat = gss.mat,
                             emb.mat = embeddings.df,
                             do.norm = "L2",
                             batch.size = batch.size,
                             n.models = n.models,
                             cores = 5)
```

### Save the reference model

ref.cellmeta stores:

1. [optional] colors: for plots

2. [optional] text.pos: text annotation on the plots

3. meta.data: cell meta data (embeddings + cell type information)

```{r}
ref.cellmeta <- seu.ref@misc$data.refplot
ref.cellmeta$meta.data <- FetchData(seu.ref, vars = c(paste0("UMAP_", 1:2), "cell_type", "cell_subtype"))

reference <- list(
  "models" = list(
    "umap" = umap.model
  ),
  "genes" = list(
    "gene.sets" = top.genes, # list
    "bg.genes" = bg.genes # vector
  ),
  "ref.cellmeta" = ref.cellmeta # list for reference plot
)
```

## Map Query to Reference

### Reference mapping

```{r, fig.width=12, fig.height=5}
seu.q <- readRDS("query_pbmc_small.seurat.slim.rds")
seu.q[["RNA"]]@counts <- seu.q[["RNA"]]@data
seu.q <- ProjectSVR::MapQuery(seu.q, reference = reference, ncores = 5)

p1 <- DimPlot(seu.q, reduction = "ref.umap", group.by = "cell_type", label = T)
p2 <- DimPlot(seu.q, reduction = "ref.umap", group.by = "cell_subtype", label = T)
(p1 + p2) & ggsci::scale_color_d3("category20")
```

```{r, fig.width=18, fig.height=14}
seu.q$cell_subtype2 <- seu.q$cell_subtype

PlotProjection(seu.q, reference, split.by = "cell_subtype2", ref.color.by = "cell_subtype",
               ref.size = .5, ref.alpha = .3, query.size = 1, query.alpha = .5, n.row = 4)
```

### Label transfer

```{r, fig.width=9, fig.height=5}
seu.q <- ProjectSVR::LabelTransfer(seu.q, reference, ref.label.col = "cell_subtype")

DimPlot(seu.q, reduction = "ref.umap", group.by = "knn.pred.celltype") + 
  scale_color_manual(values = reference$ref.cellmeta$colors)
```

```{r, fig.width=8, fig.height=5}
data.stat <- table(seu.q$cell_subtype, seu.q$knn.pred.celltype)
pheatmap::pheatmap(data.stat, cluster_cols = F,
                   display_numbers = T, number_format = "%.0f",
                   number_color = "black")
```

<details>
  <summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>


